---
title: "p8105_hw6_rs4390"
author: "Rae Spriggs"
date: "2022-11-25"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(rvest)
library(modelr)
library(mgcv)
library(patchwork)
set.seed(1)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

# Problem 1

## Creating a dataframe
```{r, warning = FALSE, message = FALSE}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```

## Bootstrapping 

```{r, warning = FALSE, message = FALSE}


```


# Problem 2 

## Reading in and cleaning homicide data

```{r, warning = FALSE, message = FALSE}
homicides = read_csv("./data/homicide-data.csv")

total_homicide = homicides %>% 
 mutate(city = str_replace(city, "$", ", ")) %>% 
  mutate(city_state = 
    paste0(city, state)) %>% 
  group_by(city_state) %>% 
  mutate(unsolved = 
           case_when(disposition == 'Closed without arrest' | 
                    disposition == 'Open/No arrest' ~ TRUE, 
                    disposition == 'Closed by arrest'~ FALSE)) %>% 
  mutate(
    victim_age = as.numeric(victim_age),
    victim_race = fct_relevel(victim_race, "White"),  
    victim_sex = fct_relevel(victim_sex, "Male")) %>% 
  filter(city_state != "Dallas, TX" && 
         city_state != "Pheonix, AZ" && 
         city_state != "Kansas City, MO" && 
         city_state != "Tulsa, AL") %>% 
  filter(victim_race == "White" | victim_race == "Black") %>% 
  group_by(city_state)
```

## Fitting glm function to Baltimore 

```{r, warning = FALSE, message = FALSE}
balt_df = 
  total_homicide %>% 
  filter(city_state == "Baltimore, MD") %>% 
  select(unsolved, victim_age, victim_race, victim_sex)

fit_logistic = 
  balt_df %>%
  glm(unsolved ~ victim_age + victim_race + victim_sex, 
      data = ., 
      family = binomial()) 
  
 fit_logistic %>%  
   broom::tidy() %>% 
   mutate(
    OR = exp(estimate),
    CI_lower = exp(estimate - 1.96 * std.error),
    CI_upper = exp(estimate + 1.96 * std.error)
  ) %>% 
  select(term, OR, starts_with("CI")) %>% 
  knitr::kable(digits = 3)
```

Homicides in which the victim is Black increases odds of the homicide being unresolved by about 2x compared to victims who are White, and female victims have about a 60% reduced risk of their case remaining unsolved. There is no association between a victim's age and the odds that their homicide is unsolved. 

## Mapping the glm function for all cities

```{r, warning = FALSE, message = FALSE}
all_cities =
  total_homicide %>% 
  select(unsolved, victim_age, victim_race, victim_sex)

all_nested = 
  all_cities %>% 
  nest(data = unsolved:victim_sex)

all_unnested = 
  all_nested %>% 
  mutate(
    glm_cities = map(.x = data, ~glm(unsolved ~ victim_age + victim_race + victim_sex, 
                                     data = .x, 
                                     family = binomial())),
    results = map(glm_cities, broom::tidy)
  ) %>% 
  select(city_state, results) %>% 
  unnest(results) %>% 
  mutate(
    OR = exp(estimate),
    CI_lower = exp(estimate - 1.96 * std.error),
    CI_upper = exp(estimate + 1.96 * std.error)
  ) %>% 
  select(city_state, term, OR, starts_with("CI")) %>% 
  filter(term == "victim_sexFemale")
```

## Plot of estimated odds ratios and confidence intervals for all cities 

```{r, warning = FALSE, message = FALSE}
all_unnested %>% 
  ggplot(aes(x = reorder(city_state, -OR), y = OR)) +
  geom_point() + 
  geom_errorbar(aes(x = city_state, ymin = OR - CI_lower, ymax = OR + CI_upper)) + 
  theme(axis.text.x = element_text(angle = 65, hjust = 1, size = 6)) +
   labs(
    title = "Odds of a Homicide Being Unsolved Based on Victim Sex in 50 Major U.S. Cities",
    x = "U.S. City",
    y = "Odds Ratio")
```
On average, there is a lower odds of a homicide being unsolved if the victim is female because the odds ratios are mostly below 1. 

# Problem 3 

## Reading in and cleaning birthweight data 

```{r, warning = FALSE, message = FALSE}
birthweight = read_csv('data/birthweight.csv') %>% 
  mutate(
    babysex = recode(babysex, `1` = "male", `2` = "female"), 
    babysex = as.factor(babysex), 
    frace = recode(frace, `1` = "white", `2` = "black", `3` = "asian", `4` = "puerto rican", `8` = "other", `9` = "unknown"), 
    frace = as.factor(frace), 
    malform = recode(malform, `0` = "absent", `1` = "present"), 
    malform = as.factor(malform), 
    mrace = recode(mrace, `1` = "white", `2` = "black", `3` = "asian", `4` = "puerto rican", `8` = "other"), 
    mrace = as.factor(mrace)
  ) 
```

## Regression model for birthweight 

```{r, warning = FALSE, message = FALSE}
fit = birthweight %>% 
  mutate(
    mrace = fct_infreq(mrace),
    babysex = fct_infreq(babysex)) %>% 
    lm(bwt ~ mrace + smoken + babysex + gaweeks + parity + blength, data = .) 

fit %>% 
  broom::tidy() %>% 
  knitr::kable(digits = 4)

resid = modelr::add_residuals(birthweight, fit) %>% 
  ggplot(aes(x = bwt, y = resid)) + 
  geom_point()

pred = modelr::add_predictions(birthweight, fit) %>% 
  ggplot(aes(x = bwt, y = pred)) + 
  geom_point()

plot_of_model = resid + pred
```

When building my regression model, I included factors that are hypothesized to impact birthweight including the mother's race and smoking habits, the baby's sex, gestational age, and number of live births the mother had prior to this pregnancy. 

## 2 models for comparison

```{r}
length_gsa = birthweight %>% 
  lm(bwt ~ blength + gaweeks, data = .)

multi = birthweight %>% 
  lm(bwt ~ bhead + blength + babysex + bhead * blength + bhead * babysex + blength * babysex + bhead * blength * babysex, data = .)


birthweight %>% 
  gather_predictions(fit, length_gsa, multi) %>% 
  mutate(model = fct_inorder(model)) %>% 
  ggplot(aes(x = blength, y = bwt)) + 
  geom_point() + 
  geom_line(aes(y = pred), color = "red") + 
  facet_wrap(~model)
```

## Cross validation 

```{r}
cv_df = 
  crossv_mc(birthweight, 100) %>% 
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble)
  )

cv_mutated = 
  cv_df %>% 
  mutate(
    fit = map(train, ~lm(bwt ~ mrace + smoken + babysex + gaweeks + parity + blength, data = .x)),
    length_gsa = map(train, ~lm(bwt ~ blength + gaweeks, data = .x)),
    multi = map(train, ~lm(bwt ~ bhead + blength + babysex + bhead * blength + bhead * babysex + blength * babysex + bhead * blength * babysex, data = .x))) %>% 
  mutate(
    rmse_fit = map2_dbl(fit, test, ~rmse(model = .x, data = .y)),
    rmse_length_gsa = map2_dbl(length_gsa, test, ~rmse(model = .x, data = .y)),
    rmse_multi = map2_dbl(multi, test, ~rmse(model = .x, data = .y)))

cv_mutated %>% 
  select(starts_with("rmse")) %>% 
  pivot_longer(
    everything(),
    names_to = "model", 
    values_to = "rmse",
    names_prefix = "rmse_") %>% 
  mutate(model = fct_inorder(model)) %>% 
  ggplot(aes(x = model, y = rmse)) + geom_violin()
```
Our model seems reasonable in comparison to the other two models, with RMSEs being fairly comparable. However the multi model has the lowest RMSEs of the threee, so it is technically the best fit. 
