---
title: "p8105_hw6_rs4390"
author: "Rae Spriggs"
date: "2022-11-25"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(rvest)
library(modelr)
library(mgcv)
set.seed(1)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

# Problem 1

## Creating a dataframe
```{r cars}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```

## Bootstrapping 

```{r}


```


# Problem 2 

## Reading in and cleaning homicide data

```{r}
homicides = read_csv("./data/homicide-data.csv")

total_homicide = homicides %>% 
 mutate(city = str_replace(city, "$", ", ")) %>% 
  mutate(city_state = 
    paste0(city, state)) %>% 
  group_by(city_state) %>% 
  mutate(unsolved = 
           case_when(disposition == 'Closed without arrest' | 
                    disposition == 'Open/No arrest' ~ TRUE, 
                    disposition == 'Closed by arrest'~ FALSE)) %>% 
  mutate(
    victim_age = as.numeric(victim_age),
    victim_race = fct_relevel(victim_race, "White")) %>% 
  filter(city_state != "Dallas, TX" && 
         city_state != "Pheonix, AZ" && 
         city_state != "Kansas City, MO" && 
         city_state != "Tulsa, AL") %>% 
  filter(victim_race == "White" | victim_race == "Black") %>% 
  group_by(city_state)
```

## Fitting glm function to Baltimore 

```{r}
balt_df = 
  total_homicide %>% 
  filter(city_state == "Baltimore, MD") %>% 
  select(unsolved, victim_age, victim_race, victim_sex)

fit_logistic = 
  balt_df %>%
  glm(unsolved ~ victim_age + victim_race + victim_sex, 
      data = ., 
      family = binomial()) 
  
 fit_logistic %>%  
   broom::tidy() %>% 
   mutate(
    OR = exp(estimate),
    CI_lower = exp(estimate - 1.96 * std.error),
    CI_upper = exp(estimate + 1.96 * std.error)
  ) %>% 
  select(term, OR, starts_with("CI")) %>% 
  knitr::kable(digits = 3)
```

Homicides in which the victim is Black or male increases odds of the homicide being unresolved by about 2x compared to victims who are White or female. There is no association between a victim's age and the odds that their homicide is unresolved. 

## Mapping the glm function for all cities

```{r}
all_cities =
  total_homicide %>% 
  select(unsolved, victim_age, victim_race, victim_sex)

all_nested = 
  all_cities %>% 
  nest(data = unsolved:victim_sex)

all_unnested = 
  all_nested %>% 
  mutate(
    glm_cities = map(.x = data, ~glm(unsolved ~ victim_age + victim_race + victim_sex, 
                                     data = .x, 
                                     family = binomial())),
    results = map(glm_cities, broom::tidy)
  ) %>% 
  select(city_state, results) %>% 
  unnest(results) %>% 
  mutate(
    OR = exp(estimate),
    CI_lower = exp(estimate - 1.96 * std.error),
    CI_upper = exp(estimate + 1.96 * std.error)
  ) %>% 
  select(city_state, term, OR, starts_with("CI"))
```

## Plot of estimated odds ratios and confidence intervals for all cities 

```{r}
all_unnested %>% 
  ggplot(aes(x = reorder(city_state, -OR), y = OR)) +
  geom_point() + 
  geom_errorbar(aes(x = city_state, ymin = OR - CI_lower, ymax = OR + CI_upper)) + 
  theme(axis.text.x = element_text(angle = 65, hjust = 1, size = 6)) +
   labs(
    title = "Odds of a Homicide Being Unsolved Based on Victim Sex in 50 Major U.S. Cities",
    x = "U.S. City",
    y = "Odds Ratio")
```


# Problem 3 

## Reading in and cleaning birthweight data 
```{r}
birthweight = read_csv('data/birthweight.csv') %>% 
  mutate(
    babysex = recode(babysex, `1` = "male", `2` = "female"), 
    babysex = as.factor(babysex), 
    frace = recode(frace, `1` = "white", `2` = "black", `3` = "asian", `4` = "puerto rican", `8` = "other", `9` = "unknown"), 
    frace = as.factor(frace), 
    malform = recode(malform, `0` = "absent", `1` = "present"), 
    malform = as.factor(malform), 
    mrace = recode(mrace, `1` = "white", `2` = "black", `3` = "asian", `4` = "puerto rican", `8` = "other"), 
    mrace = as.factor(mrace)
  ) 
```

## Regression model for birthweight 

```{r}
fit = birthweight %>% 
  mutate(
    mrace = fct_infreq(mrace),
    babysex = fct_infreq(babysex)) %>% 
    lm(bwt ~ mrace + smoken + babysex + gaweeks + parity, data = .) 

fit %>% 
  broom::tidy() %>% 
  knitr::kable(digits = 4)

## Question on what the appropriate plot is - why do we need both residuals and predictions? 
modelr::add_residuals(birthweight, fit) %>% 
  ggplot(aes(x = bwt, y = resid)) + 
  geom_point()

modelr::add_predictions(birthweight, fit) %>% 
  ggplot(aes(x = bwt, y = pred)) + 
  geom_point()

```

When building my regression model, I included factors that are hypothesized to impact birthweight including the mother's race and smoking habits, the baby's sex, gestational age, and number of live births the mother had prior to this pregnancy. 

## 2 models for comparison

```{r}
length_gsa = birthweight %>% 
  lm(bwt ~ blength + gaweeks, data = .)

multi = birthweight %>% 
  lm(bwt ~ bhead + blength + babysex + bhead * blength + bhead * babysex + blength * babysex + bhead * blength * babysex, data = .)

## Attempt to get an idea of fit of our model (not sure what should be on the axes)
birthweight %>% 
  gather_predictions(fit, length_gsa, multi) %>% 
  mutate(model = fct_inorder(model)) %>% 
  ggplot(aes(x = mrace, y = bwt)) + 
  geom_point() + 
  geom_line(aes(y = pred), color = "red") + 
  facet_wrap(~model)
```

## Cross validation 

```{r}
cv_df = 
  crossv_mc(birthweight, 100) %>% 
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble)
  )

cv_mutated = 
  cv_df %>% 
  mutate(
    fit = map(train, ~lm(bwt ~ mrace + smoken + babysex + gaweeks + parity, data = .x)),
    length_gsa = map(train, ~lm(bwt ~ blength + gaweeks, data = .x)),
    multi = map(train, ~lm(bwt ~ bhead + blength + babysex + bhead * blength + bhead * babysex + blength * babysex + bhead * blength * babysex, data = .x))) %>% 
  mutate(
    rmse_fit = map2_dbl(fit, test, ~rmse(model = .x, data = .y)),
    rmse_length_gsa = map2_dbl(length_gsa, test, ~rmse(model = .x, data = .y)),
    rmse_multi = map2_dbl(multi, test, ~rmse(model = .x, data = .y)))

cv_mutated %>% 
  select(starts_with("rmse")) %>% 
  pivot_longer(
    everything(),
    names_to = "model", 
    values_to = "rmse",
    names_prefix = "rmse_") %>% 
  mutate(model = fct_inorder(model)) %>% 
  ggplot(aes(x = model, y = rmse)) + geom_violin()

## The HW problem wants a comparison based on cross-validated prediction error - I used RMSE like Jeff did in class, so will have to revisit this problem

```

